A SAT-Based Counterexample Guided Method for Unbounded Synthesis

### 2016-01-29 paper.tex
More edits to section 2

A \emph{safety game}, $G = \langle X, L_u, L_c, \delta, I, E \rangle$,
is defined over boolean state variables $X$, uncontrollable label variables $L_u$, and
controllable label variables $L_c$.  $I$ is the initial state of the game given as a 
valuation of state variables.  $E(X)$ is the set of error states represented by its 
characteristic formula.  The transition relation $\delta(X, L_u, L_c, X')$ of the game 
is a boolean formula that relates current state and label to the set of possible next 
states of the game.  We assume deterministic games, where 
$\delta(x,u,c,x'_1) \land \delta(x,u,c,x'_2) \implies (x'_1=x'_2)$.

At every round of the game, the \emph{environment} picks an
uncontrollable label, the \emph{controller} responds by choosing a controllable 
label and the game transitions into a new state according to $\delta$. 
A \emph{run} of a game $(x_0, u_0, c_0), (x_1, u_1, c_1) \dots (x_n, u_n,
c_n)$ is a chain of state and label pairs s.t.\,$\delta(x_k, u_k, c_k, x_{k+1})$. 
A run is winning for the controller if $x_0 = I \land \forall i \in \{1..n\} (\lnot E(x_i))$. 
In a \emph{bounded game} with bound $n$ all runs are restricted to length $n$, whereas unbounded 
games consider runs of infinite length. Since we consider only deterministic games, a run 
is uniquely described by a list of assignments to $L_u$ and $L_c$.

A \emph{controller strategy} $\pi^c : (X, L_u) \to L_c$ is a mapping of states
and uncontrollable inputs to controllable labels. A controller strategy is
winning in a bounded game of rank $n$ if all runs $(x_0, u_0, \pi^c(x_0, u_0)),
(x_1, u_1, \pi^c(x_1, u_1)) \dots$ $(x_n, u_n, \pi^c(x_n, u_n))$ are winning.
Bounded \emph{realizability} is the problem of determining the existence of
such a strategy for a bounded game.

An \emph{environment strategy} $\pi^e : X \to L_u$ is a mapping of states to
uncontrollable labels. A bounded run is winning for the environment if $x_0
= I \land \exists i \in \{1..n\} (E(x_i))$ and an environment strategy is
winning for a bounded game if all runs $(x_0, \pi^e(x_1), c_1), (x_1,
\pi^e(x_1), c_1) \dots (x_n, \pi^e(x_n), c_n)$ are winning for the environment.
Safety games are zero sum, therefore the existence of a winning controller strategy
implies the nonexistence of a winning environment strategy and vice versa.

@@

Realisability for a bounded safety game can be solved by checking a quantified
formula consisting of the game's transition relation \emph{unrolled} to the
length of the bounded game. The quantifiers represent the two players of the
game: universal for the controller and existential for the environment.  Just
as the players alternate in choosing actions every step of the game, the
formula is prefixed by a quantifier alternation over $L_u$ and $L_c$ for each
step.

@@

\tikzset{every node/.style={solid}}
\tikzstyle{fixed}=[solid]
\tikzstyle{unfixed}=[dash pattern = on 2pt off 2pt]
\begin{figure}
    \centering
    \begin{subfigure}[t]{.32\textwidth}
        \centering
        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
            \node [circle,draw] (root){}
                child {node [circle,draw] {}
                    child {node [circle,draw] {}
                        edge from parent [fixed] node [left] {$b$}
                    }
                    edge from parent [fixed] node [left] {$a$}
                }
                child {node [circle,draw] {}
                    edge from parent [fixed] node [right] {$d$}
                }
                node [above=4pt] {$\langle s, k \rangle$};
        \end{tikzpicture}
        \captionsetup{width=.8\textwidth}
        \caption{AGT}
        \label{fig:agt}
    \end{subfigure}%
    \begin{subfigure}[t]{.32\textwidth}
        \centering
        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
            \node [circle,draw] (root){}
                child {node [circle,draw] {}
                    child {node [circle,draw] {}
                        node [left=4pt] {$\epsilon$}
                        edge from parent [fixed] node [left] {$b$}
                    }
                    node [left=4pt] {$\beta$}
                    edge from parent [fixed] node [left] {$a$}
                }
                child {node [circle,draw] {}
                    node [left=4pt] {$\gamma$}
                    edge from parent [fixed] node [right] {$d$}
                }
                node [right=4pt] {$\alpha$}
                node [above=4pt] {$\langle s, k \rangle$};
        \end{tikzpicture}
        \captionsetup{width=.8\textwidth}
        \caption{AGT with partial strategy}
        \label{fig:strategy}
    \end{subfigure}%
    \begin{subfigure}[t]{.32\textwidth}
        \centering
        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
            \node [circle,draw] (root){}
                child {node [circle,draw] {}
                    child {node [circle,draw] {}
                        edge from parent [fixed] node [left] {$b$}
                    }
                    edge from parent [fixed] node [left] {$a$}
                }
                child {node [circle,draw] {}
                    child {node [circle,draw] {}
                        edge from parent [fixed] node [right] {$e$}
                    }
                    edge from parent [fixed] node [right] {$d$}
                }
                node [above=4pt] {$\langle s, k \rangle$};
        \end{tikzpicture}
        \captionsetup{width=.8\textwidth}
        \caption{Refined AGT}
        \label{fig:refined}
    \end{subfigure}
    \caption{Abstract game trees}
    \label{fig:alltrees}
\end{figure}

@@

Given a controller (environment)
abstract game tree $T$ a \emph{partial strategy} $Strat: \textsc{nodes}(T) \rightarrow L_u$ 
($Strat: \textsc{nodes}(T) \rightarrow L_c$) labels each node of the tree with
the opponent's action to be played in that node.   Figure~\ref{fig:strategy}
shows an example partial strategy.  The environment starts by choosing action
$\alpha$.  If the controller plays $a$, the environment responds with $\beta$
in the next round, and so on. Given a partial strategy $Strat$, we can map each
leaf $l$ of the abstract game tree to $\langle
s',i'\rangle=\textsc{outcome}(\langle s, i\rangle, Strat, l)$ obtained by
playing all controllable and uncontrollable actions on the path from the root
to the leaf.

@@

The algorithm takes a concrete game $G$ with maximum bound $n$ as an implicit
argument.  In addition, it takes a player $p$ (controller or environment),
state $s$, bound $k$ and an abstract game tree $T$ and returns a winning
partial strategy for $\textsc{opponent}(p)$, if one exists.  The initial
invocation of the algorithm takes the initial state $\langle I, n\rangle$ and
an empty abstract game tree $\emptyset$.  Initially the solver is playing on
behalf of the environment since that player takes the first move in every step.
The empty game tree does not constrain opponent moves, hence solving such an
abstraction is equivalent to solving the original concrete game.

@@

The CEGAR loop depends on the ability to guess candidate partial strategies in
\textsc{findCandidate}. For this purpose we use the heuristic that a partial
strategy may be winning if each \textsc{outcome} of the strategy can be extended to a
run of the game that is winning for the current player.  Clearly, if such a partial 
strategy does not exist then no winning partial strategy can exist for the abstract 
game tree.  
We can formulate this heuristic as a SAT query, which is constructed
recursively by $\textsc{treeFormula}$ (for the controller) or $\textsc{\textoverline{treeFormula}}$
(for the environment) in Algorithm~\ref{alg:treeFormula}. The tree is first extended to the maximum
bound with edges that are labeled with arbitrary opponent actions (Algorithm~\ref{alg:bounded}, line 14) as shown in Figure~\ref{}.  
Then each edge in the tree is encoded in CNF by making copies of the transition relation along with the 
winning condition.

%The heuristic is based on the observation that for many synthesis problems in
%the real world the final implementation will cooperate with its environment.
%The existential search of the SAT solver simulates cooperation between the two
%players.

@@

\begin{algorithm}
    \caption{Tree formulas for Controller and Environment respectively}
    \label{alg:treeFormula}
    \begin{algorithmic}[1]
        \Function{treeFormula}{$T$}
        \If{$\Call{rank}{T} = 0$}
        \State \Return{ $\lnot \Call{E}{X_{T}}$ }
        \Else
        \State \Return{$\lnot \Call{E}{X_{T}} \land$ \\
            $$\bigwedge_{n \in \Call{succ}{T}}(\Call{$\delta$}{X_n, U_n, C_n, X'_n} \land \Call{label}{n} \land \Call{treeFormula}{n})$$
        }
        \EndIf
        \EndFunction
        \algstore{tf1}
    \end{algorithmic}

    \begin{algorithmic}[1]
        \algrestore{tf1}
        \Function{\textoverline{treeFormula}}{$T$}
        \If{$\Call{rank}{T} = 0$}
        \State \Return{\Call{E}{$X_{T}$}}
        \Else
        \State \Return{ $\Call{E}{X_{T}} \lor$ \\
        $$\bigvee_{n \in \Call{succ}{T}}(\Call{$\delta$}{X_n, U_n, C_n, X'_n} \land \Call{label}{n} \land \Call{\textoverline{treeFormula}}{n})$$ }
        \EndIf
        \EndFunction
    \end{algorithmic}
\end{algorithm}

@@

Bounded synthesis can be used to prove the existence of a winning strategy for
the environment on the unbounded game by providing a witness. For the
controller, the strongest claim that can be made is that the strategy is
winning as long as the game does not extend beyond the bound.

It is possible to set a bound such that all runs in the unbounded game will be
considered. The na\"ive approach is to use size of the state space as the bound
($2^{|X|}$) so that all states may be explored by the algorithm. A more nuanced
approach is to use the diameter of the game \cite{biere1999}, which is the
smallest number $d$ such that for any state $x$ there is a path of length $\leq
d$ to all other reachable states. This approach is common in bounded model
checking but for synthesis it quickly becomes intractable to consider such long
runs.

When performing model checking or synthesis with BDDs \cite{burch1990} the set
of states that are winning for the environment is iteratively constructed by
computing the states from which the environment can force the game into the
previous winning set. Eventually this process reaches a fixed point and the
total set of environment winning states is known.

We apply a similar fixed point computation on game states to the bounded
synthesis algorithm in order to promote it to unbounded games. Our unbounded
procedure iteratively executes the bounded algorithm while recording states
from abstract game trees. By carefully managing the way states are recorded we
guarantee that a fixed point proves the controller's strategy correct.

\subsection{Extending Bounded Synthesis to Unbounded Games}

@@

The unbounded synthesis algorithm enhances the bounded method with
computational learning. States are carefully recorded from unsatisfiable
abstract games while maintaining an invariant ensuring the monotonicity of
learning.

During the CEGAR loop of the bounded algorithm, each of the competing solvers
searches for satisfying assignments of labels to abstract game trees. When the
abstract game is unsatisfiable this indicates that the states represented by
nodes in the abstract game tree are losing for the current player. We extract
these states from the game tree using interpolation.

When executing the unbounded solver, lines 19 and 20 become active in the
bounded solver. These lines call the learning procedures when the solver fails
to find a candidate for an abstract game tree. The states symbolically
represented by nodes in the tree are losing for whichever player could not find
a winning candidate and can be extracted from the tree using interpolation.

The states in an abstract game with no controller candidate are
\emph{must-losing}. The environment can always force the game into the error
states from these states.

From abstractions with no environment candidate we record the complement of
states in the tree as \emph{may-losing}. The environment cannot reach the error
state in a number of steps equal to the distance to the bottom of the tree.

We maintain a set of states for each rank up to the current bound. We maintain
an invariant over these sets via careful construction so that they are
monotonically increasing by rank. We also ensure that the environment is unable
to force play from one set to the next. Due to these invariants, when two
adjacent sets become equivalent we know that the algorithm has reached a fixed
point and the controller is winning in the unbounded game (line 6).

%%%\begin{figure}
%%%    \centering
%%%    \begin{tikzpicture}
%%%        \node [minimum size=2cm,circle,draw] (a){$B^m[n]$};
%%%        \node [minimum size=1.5cm,circle,draw,right=1.8cm of a] (b){$B^m[n-1]$};
%%%        \node [minimum size=1cm,right=1.5cm of b] (dots){$\dots$};
%%%        \node [minimum size=0.5cm,circle,draw,right=1.5cm of dots] (c){$B^m[0]$};
%%%        \draw[-latex] (a.east) -- node [above] {$\supset$} (b.west);
%%%        \draw[-latex] (b.east) -- node [above] {$\supset$} (dots);
%%%        \draw[-latex] (dots.east) -- node [above] {$\supset$} (c.west);
%%%    \end{tikzpicture}
%%%    \caption{May-losing states}
%%%    \label{fig:maylosing}
%%%\end{figure}

@@

@@

\begin{figure}
    \centering
%%%    \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
%%%        \node [circle,draw] (root){}
%%%            child {node [circle,draw] {}
%%%                child {node [circle,draw] {}
%%%                    edge from parent [fixed] node [left] {$b$}
%%%                }
%%%                edge from parent [fixed] node [left] {$a$}
%%%            }
%%%            child {node [circle,draw] {}
%%%                edge from parent [fixed] node [right] {$d$}
%%%            }
%%%            node [above=4pt] {$\langle s, k \rangle$};
%%%    \end{tikzpicture}
    \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
        \node [] (root){}
        child {node [circle,draw] {}
                child {node [circle,draw] {}
                    edge from parent [fixed] node [left] {$a$}
                }
                child {node [circle,draw] {} edge from parent[unfixed]
                    edge from parent [fixed] node [right] {$b$}
                }
            };
    \end{tikzpicture}
%%%    \begin{tikzpicture}[overlay]
%%%        \draw [rotate = -20] (-1.9,0.35) ellipse (0.7 and 1.3);
%%%    \end{tikzpicture}
    \caption{A controller-losing game tree}
    \label{fig:interpolatetree}
\end{figure}

Consider the snippet of a game tree in Figure \ref{fig:interpolatetree}. The
tree is losing for the controller, the topmost node is at rank 1, and the $a$ and
$b$ nodes are at rank 0. Since the tree is controller-losing we know that at
least one state in the tree will uncontrollably lead to the error state.  As a
result, $\textsc{treeFormula}(T)$ is unsatisfiable. If we take a step of the
tree, such as the step where the environment plays $a$, and cut it from the
rest of the tree then $\textsc{treeFormula}(step) \land
\textsc{treeFormula}(parent)$ must too be unsatisfiable.

>>>>

A \emph{safety game}, $G = \langle X, U, C, \delta, I, E \rangle$, is defined
over boolean state variables $X$, uncontrollable action variables $U$, and
controllable action variables $C$.  $I$ is the initial state of the game given
as a valuation of state variables.  $E(X)$ is the set of error states
represented by its characteristic formula.  The transition relation $\delta(X,
U, C, X')$ of the game is a boolean formula that relates current state and
action to the set of possible next states of the game.  We assume deterministic
games, where $\delta(x,u,c,x'_1) \land \delta(x,u,c,x'_2) \implies
(x'_1=x'_2)$.

At every round of the game, the \emph{environment} picks an uncontrollable
action, the \emph{controller} responds by choosing a controllable action and
the game transitions into a new state according to $\delta$.  A \emph{run} of a
game $(x_0, u_0, c_0), (x_1, u_1, c_1) \dots (x_n, u_n, c_n)$ is a chain of
state and action pairs s.t.\,$\delta(x_k, u_k, c_k, x_{k+1})$.  A run is
winning for the controller if $x_0 = I \land \forall i \in \{1..n\} (\lnot
E(x_i))$.  In a \emph{bounded game} with bound $\kappa$ all runs are restricted
to length $\kappa$, whereas unbounded games consider runs of infinite length.
Since we consider only deterministic games, a run is uniquely described by a
list of assignments to $U$ and $C$.

A \emph{controller strategy} $\pi^c : (X, U) \to C$ is a mapping of states and
uncontrollable inputs to controllable actions. A controller strategy is winning
in a bounded game of rank $n$ if all runs $(x_0, u_0, \pi^c(x_0, u_0)), (x_1,
u_1, \pi^c(x_1, u_1)) \dots$ $(x_n, u_n, \pi^c(x_n, u_n))$ are winning.
Bounded \emph{realizability} is the problem of determining the existence of
such a strategy for a bounded game.

An \emph{environment strategy} $\pi^e : X \to U$ is a mapping of states to
uncontrollable actions. A bounded run is winning for the environment if $x_0 =
I \land \exists i \in \{1..n\} (E(x_i))$ and an environment strategy is winning
for a bounded game if all runs $(x_0, \pi^e(x_1), c_1), (x_1, \pi^e(x_1), c_1)
\dots (x_n, \pi^e(x_n), c_n)$ are winning for the environment.  Safety games
are zero sum, therefore the existence of a winning controller strategy implies
the nonexistence of a winning environment strategy and vice versa.

@@

Realisability for a bounded safety game can be solved by checking a quantified
formula consisting of the game's transition relation \emph{unrolled} to the
length of the bounded game. The quantifiers represent the two players of the
game: universal for the controller and existential for the environment.  Just
as the players alternate in choosing actions every step of the game, the
formula is prefixed by a quantifier alternation over $U$ and $C$ for each
step.

@@

\tikzset{every node/.style={solid}}
\tikzstyle{fixed}=[solid]
\tikzstyle{unfixed}=[dash pattern = on 2pt off 2pt]
\begin{figure}
    \centering
    \begin{subfigure}[t]{.25\textwidth}
        \centering
        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
            \node [circle,draw] (root){}
                child {node [circle,draw] {}
                    child {node [circle,draw] {}
                        edge from parent [fixed] node [left] {$b$}
                    }
                    edge from parent [fixed] node [left] {$a$}
                }
                child {node [circle,draw] {}
                    edge from parent [fixed] node [right] {$d$}
                }
                node [above=4pt] {$\langle s, k \rangle$};
        \end{tikzpicture}
        \captionsetup{width=.8\textwidth}
        \caption{AGT}
        \label{fig:agt}
    \end{subfigure}%
    \begin{subfigure}[t]{.25\textwidth}
        \centering
        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
            \node [circle,draw] (root){}
                child {node [circle,draw] {}
                    child {node [circle,draw] {}
                        node [left=4pt] {$\epsilon$}
                        edge from parent [fixed] node [left] {$b$}
                    }
                    node [left=4pt] {$\beta$}
                    edge from parent [fixed] node [left] {$a$}
                }
                child {node [circle,draw] {}
                    node [left=4pt] {$\gamma$}
                    edge from parent [fixed] node [right] {$d$}
                }
                node [right=4pt] {$\alpha$}
                node [above=4pt] {$\langle s, k \rangle$};
        \end{tikzpicture}
        \captionsetup{width=.8\textwidth}
        \caption{AGT with partial strategy}
        \label{fig:strategy}
    \end{subfigure}%
    \begin{subfigure}[t]{.25\textwidth}
        \centering
        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
            \node [circle,draw] (root){}
                child {node [circle,draw] {}
                    child {node [circle,draw] {}
                        edge from parent [fixed] node [left] {$b$}
                    }
                    edge from parent [fixed] node [left] {$a$}
                }
                child {node [circle,draw] {}
                    child {node [circle,draw] {}
                        edge from parent [fixed] node [right] {$e$}
                    }
                    edge from parent [fixed] node [right] {$d$}
                }
                node [above=4pt] {$\langle s, k \rangle$};
        \end{tikzpicture}
        \captionsetup{width=.8\textwidth}
        \caption{Refined AGT}
        \label{fig:refined}
    \end{subfigure}%
    \begin{subfigure}[t]{.25\textwidth}
        \centering
        \begin{tikzpicture}[scale=0.75,dash pattern = on 2pt off 2pt, level distance = 10mm]
            \node [circle,draw] (root){}
                child {node [circle,draw] {}
                    child {node [circle,draw] {}
                        child {node [circle,draw] {}
                            edge from parent [fixed] node [left] {$c$}
                        }
                        edge from parent [fixed] node [left] {$b$}
                    }
                    edge from parent [fixed] node [left] {$a$}
                }
                child {node [circle,draw] {}
                    child {node [circle,draw] {}
                        child {node [circle,draw] {}
                            edge from parent [fixed] node [left] {$f$}
                        }
                        edge from parent [fixed] node [right] {$e$}
                    }
                    edge from parent [fixed] node [right] {$d$}
                }
                node [above=4pt] {$\langle s, 3 \rangle$};
        \end{tikzpicture}
        \captionsetup{width=.8\textwidth}
        \caption{Extended AGT}
        \label{fig:extended}
    \end{subfigure}
    \caption{Abstract game trees}
    \label{fig:alltrees}
\end{figure}

@@

Given a controller (environment)
abstract game tree $T$ a \emph{partial strategy} $Strat: \textsc{nodes}(T) \rightarrow U$ 
($Strat: \textsc{nodes}(T) \rightarrow C$) labels each node of the tree with
the opponent's action to be played in that node.   Figure~\ref{fig:strategy}
shows an example partial strategy.  The environment starts by choosing action
$\alpha$.  If the controller plays $a$, the environment responds with $\beta$
in the next round, and so on. Given a partial strategy $Strat$, we can map each
leaf $l$ of the abstract game tree to $\langle
s',i'\rangle=\textsc{outcome}(\langle s, i\rangle, Strat, l)$ obtained by
playing all controllable and uncontrollable actions on the path from the root
to the leaf.

@@

The algorithm takes a concrete game $G$ with maximum bound $\kappa$ as an implicit
argument.  In addition, it takes a player $p$ (controller or environment),
state $s$, bound $k$ and an abstract game tree $T$ and returns a winning
partial strategy for $\textsc{opponent}(p)$, if one exists.  The initial
invocation of the algorithm takes the initial state $I$, bound $\kappa$ and an
empty abstract game tree $\emptyset$.  Initially the solver is playing on
behalf of the environment since that player takes the first move in every step.
The empty game tree does not constrain opponent moves, hence solving such an
abstraction is equivalent to solving the original concrete game.

@@

The CEGAR loop depends on the ability to guess candidate partial strategies in
\textsc{findCandidate}. For this purpose we use the heuristic that a partial
strategy may be winning if each \textsc{outcome} of the strategy can be
extended to a run of the game that is winning for the current player.  Clearly,
if such a partial strategy does not exist then no winning partial strategy can
exist for the abstract game tree. We can formulate this heuristic as a SAT
query, which is constructed recursively by $\textsc{treeFormula}$ (for the
controller) or $\textsc{\textoverline{treeFormula}}$ (for the environment) in
Algorithm~\ref{alg:treeFormula}.

The tree is first extended to the maximum bound with edges that are labeled
with arbitrary opponent actions (Algorithm~\ref{alg:bounded}, line 14) as shown
in Figure~\ref{}. Then copies of state and action variables are introduced for
each node in tree and opponent action variables for each edge.  Opponent
actions are set to $\textsc{action}(e)$, the action that labels the edge of the
tree $e$. The new state and action variables are connected to the successor
nodes (\textsc{succ}) by an encoding transition relation and constrained to the
winning condition of the player.

@@

\begin{algorithm}
    \caption{Tree formulas for Controller and Environment respectively}
    \label{alg:treeFormula}
    \begin{algorithmic}[1]
        \Function{treeFormula}{$T$}
        \If{$\Call{bound}{T} = 0$}
        \State \Return{ $\lnot \Call{E}{X_{T}}$ }
        \Else
        \State \Return{$\lnot \Call{E}{X_{T}} \land$ \\
            $$\bigwedge_{\langle e, n \rangle \in \Call{succ}{T}}(\Call{$\delta$}{X_T, U_e, C_n, X_n} \land U_e = \Call{action}{e} \land \Call{treeFormula}{n})$$
        }
        \EndIf
        \EndFunction
        \algstore{tf1}
    \end{algorithmic}

    \begin{algorithmic}[1]
        \algrestore{tf1}
        \Function{\textoverline{treeFormula}}{$T$}
        \If{$\Call{bound}{T} = 0$}
        \State \Return{\Call{E}{$X_{T}$}}
        \Else
        \State \Return{ $\Call{E}{X_{T}} \lor$ \\
        $$\bigvee_{\langle e, n \rangle \in \Call{succ}{T}}(\Call{$\delta$}{X_T, U_n, C_e, X_n} \land C_e = \Call{action}{e} \land \Call{\textoverline{treeFormula}}{n})$$ }
        \EndIf
        \EndFunction
    \end{algorithmic}
\end{algorithm}

@@

Bounded synthesis can be used to prove the existence of a winning strategy for
the environment on the unbounded game by providing a witness. For the
controller, the strongest claim that can be made is that the strategy is
winning as long as the game does not extend beyond the maximum bound.

It is possible to set a maximum bound such that all runs in the unbounded game will be
considered. The na\"ive approach is to use size of the state space as the bound
($2^{|X|}$) so that all states may be explored by the algorithm. A more nuanced
approach is to use the diameter of the game \cite{biere1999}, which is the
smallest number $d$ such that for any state $x$ there is a path of length $\leq
d$ to all other reachable states. However, the diameter is difficult to
estimate and can still lead to infeasibly long games.

We instead present an approach that iteratively solves games of increasing
bound while learning states. We show that reaching a fixed point in learned
states is sufficient for completeness.

%%%\subsection{Extending Bounded Synthesis to Unbounded Games}

@@

%%%The unbounded synthesis algorithm enhances the bounded method with
%%%computational learning. States are carefully recorded from unsatisfiable
%%%abstract games while maintaining an invariant ensuring the monotonicity of
%%%learning.

%%%During the CEGAR loop of the bounded algorithm, each of the competing solvers
%%%searches for satisfying assignments of labels to abstract game trees. When the
%%%abstract game is unsatisfiable this indicates that the states represented by
%%%nodes in the abstract game tree are losing for the current player. We extract
%%%these states from the game tree using interpolation.

%%%When executing the unbounded solver, lines 19 and 20 become active in the
%%%bounded solver. These lines call the learning procedures when the solver fails
%%%to find a candidate for an abstract game tree. The states symbolically
%%%represented by nodes in the tree are losing for whichever player could not find
%%%a winning candidate and can be extracted from the tree using interpolation.

%%%The states in an abstract game with no controller candidate are
%%%\emph{must-losing}. The environment can always force the game into the error
%%%states from these states.

%%%From abstractions with no environment candidate we record the complement of
%%%states in the tree as \emph{may-losing}. The environment cannot reach the error
%%%state in a number of steps equal to the distance to the bottom of the tree.

%%%We maintain a set of states for each rank up to the current bound. We maintain
%%%an invariant over these sets via careful construction so that they are
%%%monotonically increasing by rank. We also ensure that the environment is unable
%%%to force play from one set to the next. Due to these invariants, when two
%%%adjacent sets become equivalent we know that the algorithm has reached a fixed
%%%point and the controller is winning in the unbounded game (line 6).

@@

We modify the bounded synthesis algorithm to learn a subset of states from
failed attempts to find candidate strategies.

When the abstract game is unsatisfiable this indicates that the states
represented by nodes in the abstract game tree are losing for the current
player. We extract these states from the game tree using interpolation.

@@

\begin{figure}
    \centering
    \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
        \node [circle,draw] (root){}
            child {node [circle,draw] {}
                child {node [circle,draw] {}
                    edge from parent [fixed] node [left] {$b$}
                }
                edge from parent [fixed] node [left] {$a$}
            }
            child {node [circle,draw] {}
                child {node [circle,draw] {}
                    edge from parent [fixed] node [right] {$e$}
                }
                edge from parent [fixed] node [right] {$d$}
            }
            node [above=4pt] {$\langle s, 2 \rangle$};
    \end{tikzpicture}
    \caption{A controller-losing game tree}
    \label{fig:interpolatetree}
\end{figure}

Consider the game tree in Figure \ref{fig:interpolatetree}. The
tree is losing for the controller, the topmost node is at rank 1, and the $a$ and
$b$ nodes are at rank 0. Since the tree is controller-losing we know that at
least one state in the tree will uncontrollably lead to the error state.  As a
result, $\textsc{treeFormula}(T)$ is unsatisfiable. If we take a step of the
tree, such as the step where the environment plays $a$, and cut it from the
rest of the tree then $\textsc{treeFormula}(step) \land
\textsc{treeFormula}(parent)$ must too be unsatisfiable.
